---
title: "Homework 4"
author: "PSTAT 131"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
library(tidymodels)
library(ISLR)
library(ISLR2)
library(tidyverse)
library(glmnet)
library(modeldata)
library(ggthemes)
library(janitor)
library(themis)
library(naniar)
library(xgboost)
library(ranger)
library(vip)
library(corrplot)
tidymodels_prefer()
```

## Resampling

For this assignment, we will be working with **two** of our previously used data sets -- one for classification and one for regression. For the classification problem, our goal is (once again) to predict which passengers would survive the Titanic shipwreck. For the regression problem, our goal is (also once again) to predict abalone age.

Load the data from `data/titanic.csv` and `data/abalone.csv` into *R* and refresh your memory about the variables they contain using their attached codebooks.

Make sure to change `survived` and `pclass` to factors, as before, and make sure to generate the `age` variable as `rings` + 1.5!

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

### Section 1: Regression (abalone age)

#### Question 1

Follow the instructions from [Homework 2]{.underline} to split the data set, stratifying on the outcome variable, `age`. You can choose the proportions to split the data into. Use *k*-fold cross-validation to create 5 folds from the training set.

Set up the same recipe from [Homework 2]{.underline}.
```{r}
set.seed(0217)
abalone <- read_csv("abalone.csv")

abalone <- abalone %>%
  mutate(age = rings + 1.5)

abalone_split <- initial_split(data = abalone, prop = .7, strata = NULL)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)

abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height + whole_weight + shucked_weight + viscera_weight +
                           shell_weight, data = abalone_train) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("type_"):shucked_weight) %>%  # Interact all dummy vars of type with shucked_weight
  step_interact(terms = ~ longest_shell:diameter) %>%
  step_interact(terms = ~ shucked_weight:shell_weight) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())

abalone_folds <- vfold_cv(abalone_train, v = 5, strata = NULL)
```

#### Question 2

In your own words, explain what we are doing when we perform *k*-fold cross-validation:

-   What **is** *k*-fold cross-validation?
K-fold cross-validation is a model validation technique to see how well a model performs on unseen dataset, the model trains
on k-1 folds, and the last fold the model is tested.

-   Why should we use it, rather than simply comparing our model results on the entire training set?
It's good to compare our model results like this other than the whole data set, because we don't want our model to be over-
fitted to our data and not be able to perform well when we give it another dataset. 

-   If we split the training set into two and used one of those two splits to evaluate/compare our models, what resampling method would we be using?
This would be known as the train-test split.

#### Question 3

Set up workflows for three models:

1.  *k*-nearest neighbors with the `kknn` engine, tuning `neighbors`;
2.  linear regression;
3.  elastic net **linear** regression, tuning `penalty` and `mixture`.

Use `grid_regular` to set up grids of values for all of the parameters we're tuning. Use values of `neighbors` from $1$ to $10$, the default values of penalty, and values of mixture from $0$ to $1$. Set up 10 levels of each.
```{r}
knn_model <- nearest_neighbor(mode = "regression", neighbors = tune()) %>%
  set_engine("kknn")

knn_workflow <- workflow() %>%
  add_recipe(abalone_recipe) %>%
  add_model(knn_model)

knn_grid <- grid_regular(neighbors(range = c(1, 10)), levels = 10)

lm_model <- linear_reg() %>%
  set_engine("lm")

lm_workflow <- workflow() %>%
  add_recipe(abalone_recipe) %>%
  add_model(lm_model)

enet_model <- linear_reg(penalty = tune(), mixture = tune()) %>%
  set_engine("glmnet")

enet_workflow <- workflow() %>%
  add_recipe(abalone_recipe) %>%
  add_model(enet_model)

enet_grid <- grid_regular(penalty(), mixture(range = c(0, 1)), levels = 10)
```

How many models total, **across all folds**, will we be fitting to the **abalone** **data**? To answer, think about how many folds there are, how many combinations of model parameters there are, and how many models you'll fit to each fold.
There will be 555 models total, because the knn model will have 50 models, the linear regression model will have 5, and 
the Elastic net model will have 500.

#### Question 4

Fit all the models you created in Question 3 to your folded data.

*Suggest using `tune_grid()`; see the documentation and examples included for help by running `?tune_grid`*. *You can also see the code in **Lab 4** for help with the tuning process.*
```{r}
knn_tune <- tune_grid(
  knn_workflow,
  resamples = abalone_folds,  
  grid = knn_grid,        
)

enet_tune <- tune_grid(
  enet_workflow,
  resamples = abalone_folds,  
  grid = enet_grid,        
)

lm_fit <- fit_resamples(
  lm_workflow,
  resamples = abalone_folds,  
)
```

#### Question 5

Use `collect_metrics()` to print the mean and standard errors of the performance metric ***root mean squared error (RMSE)*** for each model across folds.
```{r}
knn_metrics <- collect_metrics(knn_tune) %>%
  filter(.metric == "rmse")

enet_metrics <- collect_metrics(enet_tune) %>%
  filter(.metric == "rmse")

lm_metrics <- collect_metrics(lm_fit) %>%
  filter(.metric == "rmse")

knn_metrics
enet_metrics
lm_metrics
```

Decide which of the models has performed the best. Explain how/why you made this decision. Note that each value of the tuning parameter(s) is considered a different model; for instance, KNN with $k = 4$ is one model, KNN with $k = 2$ another.
Although it was very close I'll decide to go with the linear model, because the standard error was better than what I was 
usually seeing with results in the Elastic net model that had the same RMSE.

#### Question 6

Use `finalize_workflow()` and `fit()` to fit your chosen model to the entire **training set**.

Lastly, use `augment()` to assess the performance of your chosen model on your **testing set**. Compare your model's **testing** RMSE to its average RMSE across folds.
```{r}
final_lm_fit <- fit(lm_workflow, data = abalone_train)

lm_predictions <- augment(final_lm_fit, new_data = abalone_test)

test_rmse <- lm_predictions %>%
  metrics(truth = age, estimate = .pred) %>%
  filter(.metric == "rmse")

cv_rmse <- collect_metrics(lm_fit) %>%
  filter(.metric == "rmse")

print(paste(test_rmse))
print(paste("Cross-Validation RMSE:", cv_rmse$mean))
```

### Section 2: Classification (Titanic survival)

#### Question 7

Follow the instructions from [Homework 3]{.underline} to split the data set, stratifying on the outcome variable, `survived`. You can choose the proportions to split the data into. Use *k*-fold cross-validation to create 5 folds from the training set.
```{r}
set.seed(0217)
titanic_data <- read_csv("data/titanic.csv")
titanic_data <- titanic_data %>%
  mutate(survived = factor(survived, levels = c("Yes", "No")),
         pclass = factor(pclass))

data.titanic.split <- initial_split(titanic_data, strata = survived, prop = 0.7)

titanic_train <- training(data.titanic.split)
titanic_test <- testing(data.titanic.split)

titanic_recipe <- recipe(survived ~ pclass + sex + age + sib_sp + parch + fare, data = titanic_train) %>%
  step_impute_linear(age) %>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_interact(terms = ~ sex_female:fare) %>%
  step_interact(terms = ~ age:fare)

titanic_folds <- vfold_cv(titanic_train, v = 5, strata = NULL)
```

#### Question 8

Set up the same recipe from [Homework 3]{.underline} -- but this time, add `step_upsample()` so that there are equal proportions of the `Yes` and `No` levels (you'll need to specify the appropriate function arguments). *Note: See Lab 5 for code/tips on handling imbalanced outcomes.*
```{r}
titanic_recipe_balanced <- recipe(survived ~ pclass + sex + age + sib_sp + parch + fare, data = titanic_train) %>%
  step_impute_linear(age) %>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_interact(terms = ~ sex_female:fare) %>%
  step_interact(terms = ~ age:fare) %>%
  step_upsample(survived, over_ratio = 0.5, skip = TRUE)
```

#### Question 9

Set up workflows for three models:

1.  *k*-nearest neighbors with the `kknn` engine, tuning `neighbors`;
2.  logistic regression;
3.  elastic net **logistic** regression, tuning `penalty` and `mixture`.

Set up the grids, etc. the same way you did in Question 3. Note that you can use the same grids of parameter values without having to recreate them.
```{r}
knn_model_titanic <- nearest_neighbor(mode = "classification", neighbors = tune()) %>%
  set_engine("kknn")

lm_model_titanic <- logistic_reg() %>%
  set_engine("glm")

enet_model_titanic <- logistic_reg(penalty = tune(), mixture = tune()) %>%
  set_engine("glmnet")

knn_workflow_titanic <- workflow() %>%
  add_recipe(titanic_recipe_balanced) %>%
  add_model(knn_model_titanic)

lm_workflow_titanic <- workflow() %>%
  add_recipe(titanic_recipe_balanced) %>%
  add_model(lm_model_titanic)

enet_workflow_titanic <- workflow() %>%
  add_recipe(titanic_recipe_balanced) %>%
  add_model(enet_model_titanic)
```

#### Question 10

Fit all the models you created in Question 9 to your folded data.
```{r}
knn_tune_titanic <- tune_grid(
  knn_workflow_titanic,
  resamples = titanic_folds,
  grid = knn_grid,
)

enet_tune_titanic <- tune_grid(
  enet_workflow_titanic,
  resamples = titanic_folds,
  grid = enet_grid,
)

lm_fit_titanic <- fit_resamples(
  lm_workflow_titanic,
  resamples = titanic_folds,
)
```

#### Question 11

Use `collect_metrics()` to print the mean and standard errors of the performance metric ***area under the ROC curve*** for each model across folds.
```{r}
knn_auc_metrics <- collect_metrics(knn_tune_titanic) %>%
  filter(.metric == "roc_auc")
print(knn_auc_metrics)

enet_auc_metrics <- collect_metrics(enet_tune_titanic) %>%
  filter(.metric == "roc_auc")
print(enet_auc_metrics)

lm_auc_metrics <- collect_metrics(lm_fit_titanic) %>%
  filter(.metric == "roc_auc")
print(lm_auc_metrics)
```

Decide which of the models has performed the best. Explain how/why you made this decision.
I think the best model was the KNN-model, because it had some of the best MSRE than the other models.

#### Question 12

Use `finalize_workflow()` and `fit()` to fit your chosen model to the entire **training set**.

Lastly, use `augment()` to assess the performance of your chosen model on your **testing set**. Compare your model's **testing** ROC AUC to its average ROC AUC across folds.
```{r}
best_knn_params <- select_best(knn_tune_titanic, metric = "roc_auc")
final_knn_workflow <- finalize_workflow(knn_workflow_titanic, best_knn_params)
final_knn_fit <- fit(final_knn_workflow, data = titanic_train)

knn_predictions_titanic <- augment(final_knn_fit, new_data = titanic_test)

test_roc_auc <- roc_auc(
  data = knn_predictions_titanic,
  truth = survived,
  .pred_Yes
)

knn_cv_auc <- collect_metrics(knn_tune_titanic) %>%
  filter(.metric == "roc_auc")
print(knn_cv_auc)
print(test_roc_auc)
```